---
title: "HW 1"
author: "Adon Rosen"
date: 'Date: `r Sys.Date()`'
output: pdf_document
---
  
```{r, echo=F, message=F, warning=FALSE}
library(knitr)
opts_knit$set(root.dir='~/Documents/Psych5013/HW1')  #Don't combine this call with any other chunk -especially one that uses file paths.
```


<!-- Set the report-wide options, and point to the external code file. -->
```{r set_options, echo=F, warning=FALSE}
# cat("Working directory: ", getwd())
report_render_start_time <- Sys.time()
opts_chunk$set(
  results    = 'show',
  comment    = NA,
  tidy       = FALSE,
  fig.width  = 10,
  fig.height = 6,
  fig.path   = 'figure-png/'
)
# echoChunks <- FALSE
options(width=80) #So the output is 50% wider than the default.
read_chunk("./homeWork1.R") #This allows knitr to call chunks tagged in the underlying *.R file.
```

```{r load-packages, echo=FALSE, results='show', message=FALSE, warning=FALSE}
```

```{r load-data, echo=FALSE, results='show', message=FALSE, warning=FALSE}
```

# Question 3
## a.
```{r graph-model-1, echo=FALSE, results='show', warning=FALSE, message=FALSE}
```

## b.
The y intercept here indicates what a predicted crime rate would be given a poverty rate of 0. Specifically, if there was no poverty rate, then we would expect a violent crime rate of 209.9.

## c.
Predicted value = 209.9 + 25.5 * 10.7 ; this is equal to `r 209.9 + 25.5 * 10.7`

## d.
The residual is equal to Y - $\hat{Y}$; here Y = 805 and $\hat{Y}$ = `r 209.9 + 25.5 * 10.7`. So the residual is equal to `r 805 - (209.9 + 25.5 * 10.7)`. This residual indicates we underpredicted the observed value for the state of Massachusetts.

## e.
If two states differ by 10 in their poverty rates they will differ by 255 in their predicted violent crime rate.

## f.
First find the minimum value: `r 209.9 + 25.5 * 8`

Now find the maximum value: `r 209.9 + 25.5 * 24.7`

Now find the differences to obtain the range of predicted values: `r (209.9 + 25.5 * 24.7) - (209.9 + 25.5 * 8)`

## g.
The pearson correlation between poverty rate and violent crime rate is positive. As poverty rate increases, so does violent crime.

# Question 4
## a. 
This equation is not realistic, because the data would quickly reach the ceiling of possible college GPA values. For instance a student with a high school GPA of .5 would be predicted to have a college GPA of 4.0, and anyone with a GPA above .5 would be expected to have a GPA greater than 4.0, which is not possible on the provided scale.

## b.
The slope here indicates that a 1 unit increase in high school GPA yields a .7 increase in an estimated college GPA.

## c.
1. $\hat{Y}_3$ = `r .5 + .7*3`
1. $\hat{Y}_4$ = `r .5 + .7*4`

## d.
Here the y-intercept is = 0; and the slope is 1. This indicates that there is an identical relationship in high school GPA and predicted college GPA.

# Question 10
## a.
$\hat{Y} = -37.229 + 42.969{\beta}\_{bedrooms}$

The slope here is 42.969 indicating for every additional bedroom the houses estimated price will increase by 42.969 thousands. THis is a positive relationship such that as one increases so does the other.

## b. 
1. $\hat{Y}_2$ = `r -37.229 + 42.969*2`
1. $\hat{Y}_3$ = `r -37.229 + 42.969*3`
1. $\hat{Y}_4$ = `r -37.229 + 42.969*4`

## c.
The $\hat{Y}_3$ = `r -37.229 + 42.969*3` whereas the observed value is 48.5. The residual is equal to $Y - \hat{Y}$ which equals to `r 48.5 - (-37.229 + 42.969*3)`

## d. 
In order to calculate the corellation from the slope first multiply the slope of the regression line by the standard deviation of X and then divide by the standard deviation of Y.

Slope: 42.969

SDx: 0.6070

SDy: 44.1841

Formula: (Slope * SDx) / SDy = `r (42.969 * 0.6070) / 44.1841`

## e.
The coefficient of determination is: 0.3484 ; this means that our model can explain roughly 34% of the variation observed in the outcome

## f.
** I need to finish this question later**



# Question 12
## a.






# Question 35
The answer is B