---
title: "Final"
author: "Adon Rosen"
date: 'Date: `r Sys.Date()`'
output: pdf_document
---
  
```{r, echo=F, message=F, warning=FALSE}
library(knitr)
opts_knit$set(root.dir='./')  #Don't combine this call with any other chunk -especially one that uses file paths.
```


<!-- Set the report-wide options, and point to the external code file. -->
```{r set_options, echo=F, warning=FALSE}
# cat("Working directory: ", getwd())
report_render_start_time <- Sys.time()
opts_chunk$set(
  results    = 'show',
  comment    = NA,
  tidy       = FALSE,
  fig.width  = 10,
  fig.height = 6,
  fig.path   = 'figure-png/'
)
# echoChunks <- FALSE
options(width=80) #So the output is 50% wider than the default.
read_chunk("./Final.R") #This allows knitr to call chunks tagged in the underlying *.R file.
```


```{r load-packages, echo=FALSE, results='show', message=FALSE, warning=FALSE}
```
```{r load-data, echo=FALSE, results='show', message=FALSE, warning=FALSE}
```
```{r declare-functions, echo=FALSE, results='show', message=FALSE, warning=FALSE}
```


# Question 2

## a.

The numerator degrees of freedom is: 19

The denonminator degrees of freedom is: 3

KJHFIUEHWEKJENHFIOUHF(RY)

## b. 

The n for this model was: 26

## c. 

The root MSE is: 10.28945

This value reflects the conditional standard deviation, or the standard deviation of the residual values.

## d.

The dependent variables mean value is: 61.34783

## e.

The best estimate for the PRE is 0.6210; this is the PRE when taking in to account the number of variables used to train the model.

## f.

The value of the intercept term is 162.87590

This value reflects the predicted outcome when all independent variables are 0.

## g.

The variable with the greatest predictive influence when accounting for covariates is x3

This can be observed in various ways, perhaps the most convincing is the t value, which is the largest t value for any predictor other than the intercept, furthermore the standardaized coefficeint is also the largest, suggesting the strongest conditioned relationship exists between the DV and x3.

## h.

A parameter estimate reflects the change in the DV in one unit change of the IV.

For instance, using the raw coefficients, as x3 changes from 3 to 4; there is a decrease in y of -1.21032 units.

## i. 

# Question 3

Model DF: 3

Model Sum of Squares: `r (9.1226852 * qf(0.998, df1=3, df2=18)) * 3`

Model Mean Square: `r 9.1226852 * qf(0.998, df1=3, df2=18)`

Model F value: `r qf(0.998, df1=3, df2=18)`

Error DF: 18

Error Sum of Squares: `r 489.8181818 - (9.1226852 * qf(0.998, df1=3, df2=18))`

Corrected Total DF: 21

Root MSE: `r sqrt(9.1226852 * qf(0.998, df1=3, df2=18))`

### Type 1 Anova Table

Prof DF: 1

Prof Type I SS: `r 27.65 * 9.1226852`

Prof Mean Square: `r 27.65 * 9.1226852`

Sex DF: 1

Sex Type I SS: `r 7.98 * 9.1226852`

Sex Mean Square: `r 7.98 * 9.1226852`

PROF$*$SEX DF: 1

PROF$*$SEX Type I SS: `r qf(.2049, df1=3, df2=18) * 9.1226852`

PROF$*$SEX Mean Square: `r qf(.2049, df1=3, df2=18) * 9.1226852`

PROF$*$SEX F Value: `r qf(.2049, df1=3, df2=18)`

### Type II Anova Table

PROF DF: 1

SEX DF: 1

PROF$*$SEX DF: 1

PROF$*$SEX Type II SS: 0.6337719

PROF$*$SEX F Value: `r qf(.2049, df1=3, df2=18)`

### Type III Anova Table

PROF DF: 1

SEX DF: 1

PROF$*$SEX DF: 1

PROF$*$SEX Type III SS: `r qf(.2049, df1=3, df2=18) * 9.1226852`

PROF$*$SEX Mean Square: `r qf(.2049, df1=3, df2=18) * 9.1226852`

PROF$*$SEX F Value: `r qf(.2049, df1=3, df2=18)`