---
title: "Quiz 2"
author: "Adon Rosen"
date: 'Date: `r Sys.Date()`'
output: pdf_document
---
  
```{r, echo=F, message=F, warning=FALSE}
library(knitr)
opts_knit$set(root.dir='./')  #Don't combine this call with any other chunk -especially one that uses file paths.
```


<!-- Set the report-wide options, and point to the external code file. -->
```{r set_options, echo=F, warning=FALSE}
# cat("Working directory: ", getwd())
report_render_start_time <- Sys.time()
opts_chunk$set(
  results    = 'show',
  comment    = NA,
  tidy       = FALSE,
  fig.width  = 10,
  fig.height = 6,
  fig.path   = 'figure-png/'
)
# echoChunks <- FALSE
options(width=80) #So the output is 50% wider than the default.
read_chunk("./Quiz3.R") #This allows knitr to call chunks tagged in the underlying *.R file.
```


```{r load-packages, echo=FALSE, results='show', message=FALSE, warning=FALSE}
```
```{r load-data, echo=FALSE, results='show', message=FALSE, warning=FALSE}
```
```{r declare-functions, echo=FALSE, results='show', message=FALSE, warning=FALSE}
```



# Question 1
 (29 points) The following data consists of a 3 (age) x 2 (sex) natural design in which the proportion 
of pretend play between parents and infants changes as a function of age.  The DV (Y) is the observed 
proportion of parent-infant play that consists of pretend play.  The first independent variable (X1) is age, 
where 1 = 7-month old infants; 2 = 10-month old infants; and 3 = 13-month-old infants.   The second 
independent variable (X2) is sex, where 1=girls and 2=boys.   There are 8 infants in each cell of the 
design.  

## a.
```{r q-1-a, echo=FALSE, results='show', message=FALSE, warning=FALSE}
```

## b.
The unbiased RMSE is: `r summary(model.out)$sigma`

The biased RMSE is: `r rmse(model.out)`

## c. 
```{r q-1-c, echo=FALSE, results='show', message=FALSE, warning=FALSE}
```
The F value is: `r f.val.age`

The p value is: `r p.val.age`

The models being comapred are:

model full: $Y_{playtime}= $

model restricted:

## d. 
```{r q-1-d, echo=FALSE, results='show', message=FALSE, warning=FALSE}
```

The F value is: `r f.val.sex`

The p value is: `r p.val.sex`

## e.
The models being comapred are

model full: $Y_{playtime}= \beta_{0} + \beta_{sex}\times X_{i,1}+\beta_{age}\times X_{i,2} + \beta_{sex\times age}\times X_{i,3} + \epsilon_i$
model restricted: $Y_{playtime}= \beta_{0} + \beta_{sex}\times X_{i,1}+\beta_{age}\times X_{i,2} + \epsilon_i$

## f.
While it is often inappropriate to drive model selection via significance in predictors (for instance, sex may be proven to be a signficiant predictor in the literature), the data suggest here that the most parasimounious model may be: $Y_{playtime}= \beta_{0} + \beta_{age}\times X_{i,2} +  \epsilon_i$

## g.
